---
title: "Movie Industry"
output: github_document
---
data is from kaggle: https://www.kaggle.com/datasets/danielgrijalvas/movies

# Cleaning
load packages
```{r}
library(tidyverse)
library(skimr)
library(here)
library(janitor)
```

import dataset
```{r}
movies <- read_csv("~/Downloads/movies.csv")
```

review dataset
```{r}
skim_without_charts(movies)
```

remove duplicates
```{r}
movies <- distinct(movies)
```

extract the country from the "released" column
```{r}
movies <- movies %>% 
  mutate(temp = strsplit(released, split = "\\("))
movies <- cbind(movies, t(data.frame(movies$temp)))
names(movies)[names(movies) == "1"] <- "release_date"
names(movies)[names(movies) == "2"] <- "release_country"
rownames(movies) <- c(1:nrow(movies))
movies <- movies[, !names(movies) == "temp"]
movies$release_country <- str_sub(movies$release_country, 1, nchar(movies$release_country)-1)

head(movies)
```

Change data type of release_date
```{r}
movies$release_date <- gsub(",", "", movies$release_date)
movies$release_date <- as.Date(movies$release_date, "%B %d %Y")
```


remove rows with NA values
```{r}
movies <- drop_na(movies)
```

I noticed some discrepencies between "year" and the year listed in "release_date". So, let's add a new column with just the year listed from "release_date".
```{r}
movies <- movies %>% 
mutate(release_year = format(as.Date(movies$release_date), "%Y"))

head(movies)
```

# Quick Analysis
### Plots
```{r}
ggplot(data = movies) +
  geom_point(mapping = aes(x= budget, y = gross)) +
  geom_smooth(mapping = aes(x= budget, y= gross)) +
  labs(title = "Budget Vs Gross")
```

```{r}
ggplot(data = movies) +
  geom_point(mapping = aes(x= score, y = gross)) +
  geom_smooth(mapping = aes(x= score, y= gross)) +
  labs(title = "Score Vs Gross")
```

```{r}
ggplot(data = movies) +
  geom_point(mapping = (aes(x= rating, y = gross))) +
  labs(title = "Rating Vs Gross")
```

```{r}
ggplot(data = movies) +
  geom_point(mapping = aes(x= genre, y = gross)) +
  theme(axis.text.x = element_text(angle = 40)) +
  labs(title = "Genre Vs Gross")
```

```{r}
ggplot(data = movies) +
  geom_point(mapping = aes(x= votes, y = gross)) +
  geom_smooth(mapping = aes(x= votes, y= gross)) +
  labs(title = "Votes Vs Gross")
```

```{r}
ggplot(data = movies) +
  geom_point(mapping = aes(x= runtime, y = gross)) +
  geom_smooth(mapping = aes(x= runtime, y= gross)) +
  labs(title = "Runtime Vs Gross")
```

### Tables
companies with the overall highest gross
```{r}
companies <- movies %>% 
  group_by(company) %>% 
  summarise(total_gross = sum(gross))
```
```{r}
companies <- companies[order(desc(companies$total_gross)), ] %>% 
  top_n(20)

print(companies)
```

directors with the overall highest gross
```{r}
directors <- movies %>% 
  group_by(director) %>% 
  summarise(total_gross = sum(gross))
```
```{r}
directors <- directors[order(desc(directors$total_gross)), ] %>% 
  top_n(20)

print(directors)
```

writers with the overall highest gross
```{r}
writers <- movies %>% 
  group_by(writer) %>% 
  summarise(total_gross = sum(gross))
```
```{r}
writers <- writers[order(desc(writers$total_gross)), ] %>% 
  top_n(20)

print(writers)
```

stars with the overall highest gross
```{r}
stars <- movies %>% 
  group_by(star) %>% 
  summarise(total_gross = sum(gross))
```
```{r}
stars <- stars[order(desc(stars$total_gross)), ] %>% 
  top_n(20)

print(stars)
```
